---
title: "project4"
author: "Yushi Pan, yp2560"
date: "3/28/2021"
output: pdf_document
---

```{r}
library(dplyr)
# Y:outcome
# A:binary treatment indicator
# V1-V185 = covariates
highDim_data = read.csv('../data/highDim_dataset.csv')
lowDim_data = read.csv('../data/lowDim_dataset.csv')

#head(highDim_data)
#head(lowDim_data)
#dim(highDim_data)
#dim(lowDim_data)
```


### algorithm: Doubly Robust Estimation + propensity score estimation: logistic regression

#### Introduction of the algorithm

* The Doubly Robust Estimation has the smallest asymptotic variance. It remains consistent if the outcome models are wrong but the propensity model is right, or if the propensity model is wrong but the outcome models are right.

* Doubly Robust Estimator formula:
$$\hat{\Delta_{DR}} = N^{-1}\sum_{i=1}^{N} \frac{T_iY_i-(T_i-\hat{e_i}) \hat{m_1}(X_i)}{\hat{e_i}} - N^{-1}\sum_{i=1}^{N} \frac{(1-T_i)Y_i-(T_i-\hat{e_i}) \hat{m_0}(X_i)}{1-\hat{e_i}}$$

where $\hat{e_i}$ is the estimated propensity score for individual $i$, $\hat{m_t}(X)$ is a consistent estimate for $E(Y|T=t,X)$ and is usually obtained by regressing the observed response Y on X in group t.



```{r doubly robust estimation}
DoublyRobustEst <- function(data) {
  
  X = data %>% select(-Y, -A)
  n <- dim(data)[1]
  
  start_time1 <- Sys.time()
  #get propensity scores by using logistic regression
  logit_model <- glm(A ~., data=data[,-1], family="binomial")
  propensity <- predict(logit_model, X, type="response")
  data$ps <- propensity
  end_time1 <- Sys.time()
  
  start_time2 <- Sys.time()
  #split treatment and control group, and do regression for each group
  control <- data[data$A == 0, -2]
  treatment <- data[data$A == 1, -2]

  control_model <- lm(Y ~., data=control)
  treatment_model <- lm(Y ~., data=treatment)
  
  data$m0 <- predict(control_model, data[,-c(1,2)])
  data$m1 <- predict(treatment_model, data[,-c(1,2)])
  
  #calculate ATE
  ATE <- sum((data$A*data$Y-(data$A-data$ps)*data$m1)/data$ps)/n - 
         sum(((1-data$A)*data$Y+(data$A-data$ps)*data$m0)/(1-data$ps))/n
  
  end_time2 <- Sys.time()
  
  time1 <- end_time1 - start_time1
  time2 <- end_time2 - start_time2
  
  df <- data.frame(cbind(ATE, time1, time2))
  
  return(df)
}

```

#### lowDim data

```{r lowDim data}
DRE.low <- DoublyRobustEst(lowDim_data)
true.ATE.low <- 2.0901
diff.low <- true.ATE.low-DRE.low[,1]

cat("ATE using Doubly Robust Estimation for lowDim data = ", DRE.low[,1], "\n")
cat("The difference between the calculated ATE and the true ATE = ", diff.low, "\n")
cat("Processing time for calculating the propensity score is" ,DRE.low[,2], "seconds.\n")
cat("Processing time for the algorithm for low dimensional data is" ,DRE.low[,3], "seconds.\n")
```


#### highDim data

```{r highDim data}
DRE.high <- DoublyRobustEst(highDim_data)
true.ATE.high <- -54.8558 
diff.high <- true.ATE.high-DRE.high[,1]

cat("ATE using Doubly Robust Estimation for highDim data = ", DRE.high[,1], "\n")
cat("The difference between the calculated ATE and the true ATE = ", diff.high, "\n")
cat("Processing time for calculating the propensity score is" ,DRE.high[,2], "seconds.\n")
cat("Processing time for the algorithm for high dimensional data is" ,DRE.high[,3], "seconds.\n")
``` 



