---
title: "pairing 5"
author: "Yuqi Xing"
date: "3/29/2021"
output:
  html_document: default
  pdf_document: default
---
 

```{r results = FALSE, include=FALSE}
if(!require(MatchIt)) install.packages("MatchIt", repos = "http://cran.us.r-project.org")
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) install.packages("ggplot2", repos = "http://cran.us.r-project.org")
if(!require(lmtest)) install.packages("lmtest", repos = "http://cran.us.r-project.org")
if(!require(sandwich)) install.packages("sandwich", repos = "http://cran.us.r-project.org")

library(MatchIt)
library(dplyr)
library(ggplot2)
library(lmtest) 
library(sandwich) 
```

```{r Difference in Means Pre-Analysis}
# Reading in the datasets as R Dataframes
getwd()
setwd('~/Desktop/Spring2021-Project4-group-6-1/data')
highDim_data = read.csv('highDim_dataset.csv')
lowDim_data = read.csv('lowDim_dataset.csv')

highDim_data %>% 
  group_by(A) %>% 
  summarise(n=n(),
            mean_outcome=mean(Y),
            std_error = sd(Y) / sqrt(n))

with(highDim_data, t.test(Y ~ A))

lowDim_data %>% 
  group_by(A) %>% 
  summarise(n=n(),
            mean_outcome=mean(Y),
            std_error = sd(Y) / sqrt(n))

with(lowDim_data, t.test(Y ~ A))
```

```{r Propensity Score Estimation}
# High dimension dataset
ps_hd_est <- glm(A ~ . -Y,
                 family=binomial(), 
                 data=highDim_data)

ps_hd_df <- data.frame(prop_score = predict(ps_hd_est, type = "response"),
                       treatment = ps_hd_est$model$A)
ps_hd_df %>%
  group_by(treatment) %>% 
  summarise(mean_prop_score=mean(prop_score))
  
labs <- paste("Actual treatment indicator:", c("Treated", "Control"))
ps_hd_df %>%
  mutate(treatment = ifelse(treatment == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = prop_score)) +
  geom_histogram(color = "white") +
  facet_wrap(~treatment) +
  xlab("Probability of being exposed") +
  theme_bw()

# Low dimension dataset
ps_ld_est <- glm(A ~ . -Y,
                 family=binomial(), 
                 data=lowDim_data)

ps_ld_df <- data.frame(prop_score = predict(ps_ld_est, type = "response"),
                       treatment = ps_ld_est$model$A)
ps_ld_df %>%
  group_by(treatment) %>% 
  summarise(mean_prop_score=mean(prop_score))
  
labs <- paste("Actual treatment indicator:", c("Treated", "Control"))
ps_ld_df %>%
  mutate(treatment = ifelse(treatment == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = prop_score)) +
  geom_histogram(color = "white") +
  facet_wrap(~treatment) +
  xlab("Probability of being exposed") +
  theme_bw()
```


# Pairing 5
Algorithm: Propensity Matching

Distance Measure: Linear Propensity Score

Propensity Score Estimation: Logistic Regression
```{r Pairing 3 Implementing Matching Algo}
# Algorithm: Propensity Matching
# Distance Measure: Linear Propensity Score
# Propensity Score Estimation: Logistic Regression
system.time({
  if (length(na.omit(lowDim_data)) != length(lowDim_data)) {
    print('There are null values in the dataset')
    break
    } else {
    pair_5_low <- matchit(A ~ .-Y,
                      data=lowDim_data,
                      method="nearest",
                      distance="glm",
                      link = "linear.logit")
  }  
})



system.time({
  if (length(na.omit(highDim_data)) != length(highDim_data)) {
    print('There are null values in the dataset')
    break
    } else {
    pair_5_high <- matchit(A ~ .-Y,
                      data=highDim_data,
                      method="nearest",
                      distance="glm",
                      link = "linear.logit")
  }  
})
```

```{r Assessing Quality of Pairing 3 Matches}
pair_5_low
summary(pair_5_low, un = FALSE)
plot(pair_5_low, type = "jitter", interactive = FALSE)
plot(pair_5_low, type = 'hist')

#plot(pair_5_low)


#pair_5_high
#summary(pair_5_high, un = FALSE)
plot(pair_5_high, type = "jitter", interactive = FALSE)
plot(pair_5_high, type = 'hist')
#plot(pair_5_high)
```


```{r Preparing Data Frame based on Pairing 3 Matches}
mpair_5_low = match.data(pair_5_low)
head(mpair_5_low)

mpair_5_high = match.data(pair_5_high)
#head(mpair_5_high)
```

```{r Estimating Treatment Effects using Pairing 3 Matches}
mpair5_low_fit <- lm(Y ~ . -Y -weights -subclass,
                 data = mpair_5_low,
                 weights = weights)

coeftest(mpair5_low_fit, vcov. = vcovCL, cluster = ~subclass)


mpair5_high_fit <- lm(Y ~ . -Y -weights -subclass,
                 data = mpair_5_high,
                 weights = weights)

#coeftest(mpair5_high_fit, vcov. = vcovCL, cluster = ~subclass)
```


#Linear Propensity Score
The distance of Propensity Score is defined as:
$D_{ij}=|e_i-e_j|$
Obtained by applying the logit function on the Propensity Scores.
Matching on the linear propensity score can be particularly effective in terms of reducing bias.
Linear propensity score matching is same with propensity score to entail forming matched sets of treated and untreated subjects who share a similar value of the propensity score. Once a matched sample has been formed, the treatment effect can be estimated by directly comparing outcomes between treated and untreated subjects in the matched sample. Once the effect of treatment has been estimated in the propensity score matched sample, the variance of the estimated treatment effect and its statistical significance can be estimated. 
After the matched sets are obtained, Linear Propensity Score asks to calculate a “subclass effects” for each matched set/subclass, and then estimate overall ATE by an weighted average of the subclass effects where weights would be the number of individuals in each subclass.


Linear Propensity score performed well for low dimensional dataset as explained above for standard propensity score.
Linear Propensity Score didn't perform as well for high dimension as low dimensional dataset for the same reason discussed above for standard Propensity Score.

```{r ATE(average treatment effect), running time of pairing 3}
linear.Propensity.Score.ate <- function(data){
  start_time <- Sys.time()
  match_full<-matchit(A ~ .-Y,data=data,method="nearest",distance="glm",link = "linear.logit")
  data.fullMatching <- match.data(match_full)
  x = data.fullMatching %>% group_by(subclass,A) %>% summarise(mean_y = mean(Y), .groups = 'drop')
  group_ate = x %>% group_by(subclass) %>% summarise(treat_eff = mean_y[A == 1] - mean_y[A == 0], .groups = 'drop')
  group_n = data.fullMatching %>% group_by(subclass) %>% count()
  ate = sum(group_ate$treat_eff*group_n$n/nrow(data))
  end_time <- Sys.time()
  return(list(ATE=ate,running_time = end_time - start_time))
}

cat("ATE for low dimension is:", linear.Propensity.Score.ate(lowDim_data)$ATE, ".\n")
cat("ATE for high dimension is:", linear.Propensity.Score.ate(highDim_data)$ATE, ".\n")
cat("Processing time for low dimensional data is" ,linear.Propensity.Score.ate(lowDim_data)$running_time, "seconds", ".\n")
cat("Processing time for high dimensional data is" ,linear.Propensity.Score.ate(highDim_data)$running_time, "seconds", ".\n")


true_low_ATE<-2.0901
true_high_ATE<- -54.8558

cat("ATE error for low dimension is:", true_low_ATE-linear.Propensity.Score.ate(lowDim_data)$ATE, ".\n")
cat("ATE error for high dimension is:", true_high_ATE-linear.Propensity.Score.ate(highDim_data)$ATE, ".\n")
```































